\section*{Acknowledgement}
\addcontentsline{toc}{chapter}{Acknowledgement}
It is always a pleasure to finish a research project on a high note. It gives extra pleasure while looking back working with kind, humble and wise people. 

First of all, I would like to thank the Almighty God for helping me to successfully finish this project. Her blessings guided me through some of the hard times in my life and showed me the correct path. I would like to express my heartfelt gratitude to my supervisor Prof. Shilpa Verma for allowing me to work with her, helping me to see it through to completion. She was very supportive, encouraging, patient during the whole life-cycle of the project and the final thesis preparation. She took the pain of reading through the initial drafts of my thesis during her busy schedule and I am greatly thankful to her for the suggestions regarding the structure and content of final paper. I thank my examiner for reading through the final paper and making the necessary arrangements for the successful presentation.

Finally, I would like to thank my family, colleagues and friends, for all the support given during the project implementation stage.

\vspace*{10px}

\begin{minipage}{0.5\textwidth}
-
\end{minipage}% don't leave space after this
\begin{minipage}{0.5\textwidth}
\vspace*{55px}
\quad\quad\textbf{Rahul Bali} 16205025\\
M.Tech, Computer Science and Engineering\\
Punjab Engineering College\\
Chandigarh - 160012, India\\
Phone: +91-8901337758
Email: rahulrdb18@gmail.com 
\end{minipage}
