\chapter{Introduction}
Recommender Systems is a hot topic within Computer Science, as algorithms for the purpose of targeted advertisement get improved each year. Hence it is vital for companies to keep up with their competitors. Many recommender systems focus on the companies with business in the ecommerce sector, as shown in the review paper \cite{Park2012}. Personal integrity, or online privacy, is also a frequently discussed topic within computer science, as well as, in society and on a global level. How much and what kind of data is ethically acceptable to observe in order to produce relevant advertisements? Is it acceptable to use all data that could be retrieved from a user such as GPS-based movement, cookies, chat-history, previous purchases or other arbitrary personal data a company might have access to?

This thesis focuses on the evaluation and analysis of the performance of "Item-based Collaborative filtering" for the purpose of targeted advertisement within e-commerce. Standard algorithms are used for the evaluation and analysis. In scientific literature, a recommender system is usually interchangably used with collaborative filtering.

Comparisons are made with more standard algorithms, being used in many live recommender systems for ecommerce as of today. Whereas the algorithmic approach of hybrid algorithms have use in areas within recommender systems, very little research has been made for use within e-commerce. Recommender systems have come a long way and have the diversified in their applications and the use of techniques. Techniques used as recommendation engines are still getting better and better than the their ancestorial methods. Performance will also be analysed and evaluated taking into consideration how much personal data are observed. How much and what kind of data can be observed and manipulated for it to still be considered socially and ethically acceptable? That is while still respecting the users’ privacy. There exist research‌​ on the subject, addressing the‌​ issue of online‌​ privacy. 

\section{Background}‌​‌​‌​‌​‌​‌​‌​‌​‌​‌​‌​‌​‌​‌​‌​‌​
Matrix Factorization approach has garnered recent attention due to the Netflix Prize \cite{Bennett2007} \$1 Million contest organised in 2006. This contest gave a boost to the technology in the recommender engines for the e-commerce community. Emerging technique from the contest was Matrix Factorization and its modifications. These algorithms kept on being modified and combined with other algorithms until on September 21, 2009 \textit{BellKor's Pragmatic Chaos} team outperformed the existing algorithms by atleast 10\%. This gave birth to the advanced technique for the recommendations, today known as SVD++. This technique primarily is a latent variable technique, that means it will work as an unsupervised learning method. As for an unsupervisd techniques this method acts like an black box which has latent factors built into the matrices. These matrices are generated mathematically using the Singular Valued Decompoistion. This linear algebra technique has been used in many technologies but the use as recommendation algorithm has an vast impact on the complete e-commerce market \cite{Koren2009}.

\section{Research Objectives}
\begin{itemize}
	\item To study the parameters and performance of existing algorithms.
	\item Implement the recommender systems in the Python Language.
	\item Comparison of proposed algorithm with existing algorithms.
\end{itemize}


\section{Research Gaps}
\begin{itemize}
\item Creating the structure of Recommender System such that it is not very intrusive in
nature for user. User must feel safe while using them.
\item High amount of Dimensions cause much high number of calculations, dimensionality
reduction is a key area of interest.
\item Profiling of data on the browser, user preferences, profiling demographics, profiling,
etc.
\item Ratings must be of multiple-way input which help get the idea of real judgment of
the user.
\item Recommender system must be flexible enough to run in a way that the user wants.
\end{itemize}
